---
title: "ARIMA_GLM_shap_same"
author: "Kechen Zhao 957398"
date: "12/10/2021"
output: word_document
---

ARIMA 12months prediction modelling
```{r}
library(forecast)
library(lubridate)
library(tseries)
library(fpp)
library(ggplot2)
library(caret)
# import data
data <- read.csv("/Users/abigail/Dropbox/Data\ Science\ Project/Combined\ Datasets\ For\ AUS\ and\ US/08-19_combined.csv")
```
Construct training and test set:
```{r}
rownames(data) <- data$month
data <- subset(data, select = -month)

x <- subset(data, select = -import_ratio)
y <- subset(data, select = import_ratio)

market_share_all <- ts(y, start=c(2008.1),frequency=12)
```
```{r}
training.x <- x[1:132,]
training.y <- y[1:132,]
share.train <- ts(training.y, start=c(2008.1),frequency=12)
test.x <- x[133:144,]
test.y <- y[133:144,]
share.test <- ts(test.y, start=c(2019.1),frequency=12)
```
Since training.x and test.x are rank deficient, we will only select the significant variables for the model fitting:
```{r}
sig.x.train <- cbind(training.x$aus_beef.veal_slaughterings, training.x$aus_population_natural_increase, training.x$aus_beef.veal_exports, training.x$us_production_fedral_inspected_total.red.meat.and.poultry, training.x$aus_exports.of.beef..veal.and.live.cattle.to.US, training.x$us_livestock_commercial_mature.sheep, training.x$us_livestock_commercial_beef, training.x$us_veganism_trend_ratio, training.x$us_fed_avg_dressed_calves, training.x$us_veganism)

colnames(sig.x.train) <- c('aus_beef.veal_slaughterings', 'aus_population_natural_increase', 'aus_beef.veal_exports', 'us_production_fedral_inspected_total.red.meat.and.poultry', 'aus_exports.of.beef..veal.and.live.cattle.to.US', 'us_livestock_commercial_mature.sheep', 'us_livestock_commercial_beef', 'us_veganism_trend_ratio', 'us_fed_avg_dressed_calves', 'us_veganism')
```

```{r}
modArima <- auto.arima(share.train, xreg=sig.x.train)
summary(modArima)
```
Apply the model to test set:
```{r}
sig.x.test <- cbind(test.x$aus_beef.veal_slaughterings, test.x$aus_population_natural_increase, test.x$aus_beef.veal_exports, test.x$us_production_fedral_inspected_total.red.meat.and.poultry, test.x$aus_exports.of.beef..veal.and.live.cattle.to.US, test.x$us_livestock_commercial_mature.sheep, test.x$us_livestock_commercial_beef, test.x$us_veganism_trend_ratio, test.x$us_fed_avg_dressed_calves, test.x$us_veganism)

colnames(sig.x.test) <- c('aus_beef.veal_slaughterings', 'aus_population_natural_increase', 'aus_beef.veal_exports', 'us_production_fedral_inspected_total.red.meat.and.poultry', 'aus_exports.of.beef..veal.and.live.cattle.to.US', 'us_livestock_commercial_mature.sheep', 'us_livestock_commercial_beef', 'us_veganism_trend_ratio', 'us_fed_avg_dressed_calves', 'us_veganism')


Forecasted_values<-forecast(modArima,nrow(test.y),xreg=sig.x.test)
Final_forecasted_values<-Forecasted_values$mean
Final_forecasted_values
```
Calculating the MSE in the test dataset:
```{r}
(RMSE.ARIMA.with.Seasonality.12 <- sqrt(sum((test.y - Final_forecasted_values)^2)/12))
(MAE.ARIMA.with.Seasonality.12 <- sum(abs(test.y - Final_forecasted_values))/12)
(MAPE.ARIMA.with.Seasonality.12 <- 100*sum(abs((test.y - Final_forecasted_values)/Final_forecasted_values))/12)
```

Plotting forecast values:
```{r}
forecast(modArima,nrow(test.y),xreg=sig.x.test) -> fc
ARIMA.with.Seasonality.plot.12 <- autoplot(market_share_all, series="Data") + 
autolayer(fc, series="Forecast") + 
autolayer(fitted(fc), series="Fitted")

ARIMA.with.Seasonality.plot.12 <- ARIMA.with.Seasonality.plot.12 + labs(title = "ARIMA.with.Seasonality -- 12 Months Forecast") + labs(caption = "RMSE = 0.02501762, MAE = 0.01904903, MAPE = 9.129693")
ARIMA.with.Seasonality.plot.12
```

Plot only the foretasted period:
```{r}
forecast(modArima,nrow(test.y),xreg=sig.x.test)$mean -> fc
ARIMA.Corr.with.Seasonality.forecast.12 <- autoplot(share.test, series="Data") + autolayer(fc, series="Forecast")

ARIMA.Corr.with.Seasonality.forecast.12 <- ARIMA.Corr.with.Seasonality.forecast.12 + labs(title = "ARIMA.with.Seasonality -- 12 Months Forecast on Test set") + labs(caption = "RMSE = 0.02501762, MAE = 0.01904903, MAPE = 9.129693")
ARIMA.Corr.with.Seasonality.forecast.12
```
Model Validation: 
Jan-Dec 2014
```{r}
fit.2014.12 <- fitted(forecast(modArima,nrow(test.y),xreg=sig.x.test))[73:84]
data.2014 <- market_share_all[73:84,]
```
```{r}
(RMSE.ARIMA.with.Seasonality.12.val <- sqrt(sum((data.2014 - fit.2014.12)^2)/12))
(MAE.ARIMA.with.Seasonality.12.val <- sum(abs(data.2014 - fit.2014.12))/12)
(MAPE.ARIMA.with.Seasonality.12.val <- 100*sum(abs((data.2014 - fit.2014.12)/fit.2014.12))/12)
```
Jan-Dec 2016
```{r}
fit.2016.12 <- fitted(forecast(modArima,nrow(test.y),xreg=sig.x.test))[97:108]
data.2016 <- market_share_all[97:108,]
```
```{r}
(RMSE.ARIMA.with.Seasonality.12.val <- sqrt(sum((data.2016 - fit.2016.12)^2)/12))
(MAE.ARIMA.with.Seasonality.12.val <- sum(abs(data.2016 - fit.2016.12))/12)
(MAPE.ARIMA.with.Seasonality.12.val <- 100*sum(abs((data.2016 - fit.2016.12)/fit.2016.12))/12)
```



Fitting ARIMA without seasonality
Deseasonaling the dataset and plotting：
```{r}
decompose_data = decompose(share.train, "additive")
adjust_market_share = share.train - decompose_data$seasonal
plot(adjust_market_share)
```
Fit ARIMA to deseasoned data:
```{r}
modArima_desea <- auto.arima(adjust_market_share, xreg=sig.x.train)
modArima_desea
```
```{r}
Forecasted_values_des<-forecast(modArima_desea,nrow(test.y),xreg=sig.x.test)
Final_forecasted_values_des<-Forecasted_values_des$mean
(Final_forecasted_values_des)

(RMSE.ARIMA.without.Seasonality.12 <- sqrt(sum((test.y - Final_forecasted_values_des)^2)/12))
(MAE.ARIMA.without.Seasonality.12 <- sum(abs(test.y - Final_forecasted_values_des))/12)
(MAPE.ARIMA.without.Seasonality.12 <- 100*sum(abs((test.y - Final_forecasted_values_des)/Final_forecasted_values_des))/12)
```
Plotting forecast values after seasonality removal:
```{r}
forecast(modArima_desea,nrow(test.y),xreg=sig.x.test) -> fc1
ARIMA.without.Seasonality.plot.12 <- autoplot(market_share_all, series="Data") + 
autolayer(fc1, series="Forecast") + 
autolayer(fitted(fc1), series="Fitted")

ARIMA.without.Seasonality.plot.12 <- ARIMA.without.Seasonality.plot.12 + labs(title = "ARIMA.without.Seasonality.12") + labs(caption = "RMSE = 0.03433002, MAE = 0.02723821, MAPE = 12.09295")
ARIMA.without.Seasonality.plot.12
```
Plot only the foretasted period:
```{r}
forecast(modArima_desea,nrow(test.y),xreg=sig.x.test)$mean -> fc
ARIMA.Corr.without.Seasonality.forecast.12 <- autoplot(share.test, series="Data") + autolayer(fc, series="Forecast")

ARIMA.Corr.without.Seasonality.forecast.12 <- ARIMA.Corr.without.Seasonality.forecast.12 + labs(title = "ARIMA.Corr.without.Seasonality.forecast.12") + labs(caption = "RMSE = 0.03433002, MAE = 0.02723821, MAPE = 12.09295")
ARIMA.Corr.without.Seasonality.forecast.12
```
Use glm:
```{r}
glm.train <- glm(share.train~., data=as.data.frame(sig.x.train))
summary(glm.train)
(glm.prediction <- predict(glm.train, as.data.frame(sig.x.test)))

(RMSE.GLM.12 <- sqrt(sum((test.y - glm.prediction)^2)/12))
(MAE.GLM.12 <- sum(abs(test.y - glm.prediction))/12)
(MAPE.GLM.12 <- 100*sum(abs((test.y - glm.prediction)/glm.prediction))/12)
```

Plotting forecast values:
```{r}
fc1 <- ts(glm.prediction, start=c(2019.1),frequency=12)
fitted <- ts(glm.train$fitted.values, start=c(2008.1),frequency=12)
GLM.plot.12 <- autoplot(market_share_all, series="Data") + 
autolayer(fc1, series="Forecast") + 
autolayer(fitted, series="Fitted")

GLM.plot.12 <- GLM.plot.12 + labs(title = "GLM.12") + labs(caption = "RMSE = 0.03703912, MAE = 0.03100154, MAPE = 11.94618")
GLM.plot.12
```

Plot only the foretasted period:
```{r}
fc1 <- ts(glm.prediction, start=c(2019.1),frequency=12)
glm.corr.forecast.12 <- autoplot(share.test, series="Data") + autolayer(fc1, series="Forecast")

glm.corr.forecast.12 <- glm.corr.forecast.12 + labs(title = "glm.corr.forecast.12")+ labs(caption = "RMSE = 0.03703912, MAE = 0.03100154, MAPE = 11.94618")
glm.corr.forecast.12
```









ARIMA 6months prediction modelling
```{r}
data <- read.csv("/Users/abigail/Dropbox/Data\ Science\ Project/Combined\ Datasets\ For\ AUS\ and\ US/08-19_combined.csv")
```
Construct training and test set:
```{r}
rownames(data) <- data$month
data <- subset(data, select = -month)

x <- subset(data, select = -import_ratio)
y <- subset(data, select = import_ratio)

market_share_all <- ts(y, start=c(2008.1),frequency=12)
```
```{r}
training.x <- x[1:138,]
training.y <- y[1:138,]
share.train <- ts(training.y, start=c(2008.1),frequency=12)
test.x <- x[139:144,]
test.y <- y[139:144,]
share.test <- ts(test.y, start=c(2019.6),frequency=12)
```
Since training.x and test.x are rank deficient, we will only select the significant variables for the model fitting:
```{r}
sig.x.train <- cbind(training.x$us_livestock_commercial_lambs.and.yearlings, training.x$aus_beef.veal_exports, training.x$us_livestock_commercial_beef, training.x$aus_total.dairy.cattle.and.calves, training.x$aus_Meat.Produced....CATTLE..excl..calves.....Total..State..Original..tonnes., training.x$us_veganism, training.x$aus_avg_rainfall, training.x$us_veganism_trend_ratio, training.x$aus_Meat.Produced....CATTLE..excl..calves.....Total..State..Seasonally.adjusted..tonnes., training.x$us.cattle.numbers..1000., training.x$us_fed_avg_dressed_calves, training.x$aus_exports.of.beef..veal.and.live.cattle.to.US, training.x$us_production_fedral_inspected_total.poultry, training.x$us_cold_storage_beef, training.x$us_production_commercial_beef, training.x$us_livestock_commercial_mature.sheep, training.x$us_livestock_commercial_steers)

colnames(sig.x.train) <- c('us_livestock_commercial_lambs.and.yearlings', 'aus_beef.veal_exports', 'us_livestock_commercial_beef', 'aus_total.dairy.cattle.and.calves', 'aus_Meat.Produced....CATTLE..excl..calves.....Total..State..Original..tonnes.', 'us_veganism', 'aus_avg_rainfall', 'us_veganism_trend_ratio', 'aus_Meat.Produced....CATTLE..excl..calves.....Total..State..Seasonally.adjusted..tonnes.', 'us.cattle.numbers..1000.', 'us_fed_avg_dressed_calves', 'aus_exports.of.beef..veal.and.live.cattle.to.US', 'us_production_fedral_inspected_total.poultry', 'us_cold_storage_beef', 'us_production_commercial_beef', 'us_livestock_commercial_mature.sheep', 'us_livestock_commercial_steers')

modArima <- auto.arima(share.train, xreg=sig.x.train)
summary(modArima)
```
Apply the model to test set:
```{r}
sig.x.test <- cbind(test.x$us_livestock_commercial_lambs.and.yearlings, test.x$aus_beef.veal_exports, test.x$us_livestock_commercial_beef, test.x$aus_total.dairy.cattle.and.calves, test.x$aus_Meat.Produced....CATTLE..excl..calves.....Total..State..Original..tonnes., test.x$us_veganism, test.x$aus_avg_rainfall, test.x$us_veganism_trend_ratio, test.x$aus_Meat.Produced....CATTLE..excl..calves.....Total..State..Seasonally.adjusted..tonnes., test.x$us.cattle.numbers..1000., test.x$us_fed_avg_dressed_calves, test.x$aus_exports.of.beef..veal.and.live.cattle.to.US, test.x$us_production_fedral_inspected_total.poultry, test.x$us_cold_storage_beef, test.x$us_production_commercial_beef, test.x$us_livestock_commercial_mature.sheep, test.x$us_livestock_commercial_steers)

colnames(sig.x.test) <- c('us_livestock_commercial_lambs.and.yearlings', 'aus_beef.veal_exports', 'us_livestock_commercial_beef', 'aus_total.dairy.cattle.and.calves', 'aus_Meat.Produced....CATTLE..excl..calves.....Total..State..Original..tonnes.', 'us_veganism', 'aus_avg_rainfall', 'us_veganism_trend_ratio', 'aus_Meat.Produced....CATTLE..excl..calves.....Total..State..Seasonally.adjusted..tonnes.', 'us.cattle.numbers..1000.', 'us_fed_avg_dressed_calves', 'aus_exports.of.beef..veal.and.live.cattle.to.US', 'us_production_fedral_inspected_total.poultry', 'us_cold_storage_beef', 'us_production_commercial_beef', 'us_livestock_commercial_mature.sheep', 'us_livestock_commercial_steers')


Forecasted_values<-forecast(modArima,nrow(test.y),xreg=sig.x.test)
Final_forecasted_values<-Forecasted_values$mean
Final_forecasted_values
```
Calculating the MSE in the test dataset:
```{r}
(RMSE.ARIMA.with.Seasonality.6 <- sqrt(sum((test.y - Final_forecasted_values)^2)/6))
(MAE.ARIMA.with.Seasonality.6 <- sum(abs(test.y - Final_forecasted_values))/6)
(MAPE.ARIMA.with.Seasonality.6 <- 100*sum(abs((test.y - Final_forecasted_values)/Final_forecasted_values))/6)
```
Plotting forecast values:
```{r}
forecast(modArima,nrow(test.y),xreg=sig.x.test) -> fc
ARIMA.with.Seasonality.plot.6 <- autoplot(market_share_all, series="Data") + 
autolayer(fc, series="Forecast") + 
autolayer(fitted(fc), series="Fitted")

ARIMA.with.Seasonality.plot.6 <- ARIMA.with.Seasonality.plot.6 + labs(title = "ARIMA.with.Seasonality -- 6 Months Forecast") + labs(caption = "RMSE = 0.01438483, MAE = 0.01218461, MAPE = 4.811848")
ARIMA.with.Seasonality.plot.6
```
Plot only the foretasted period:
```{r}
forecast(modArima,nrow(test.y),xreg=sig.x.test)$mean -> fc
ARIMA.Corr.with.Seasonality.forecast.6 <- autoplot(share.test, series="Data") + autolayer(fc, series="Forecast")

ARIMA.Corr.with.Seasonality.forecast.6 <- ARIMA.Corr.with.Seasonality.forecast.6 + labs(title = "ARIMA.with.Seasonality -- 6 Months Forecast on Test Set") + labs(caption = "RMSE = 0.02518209, MAE = 0.02042195, MAPE = 9.875794")
ARIMA.Corr.with.Seasonality.forecast.6
```
Model Validation
Apr-Oct 2014:
```{r}
fit.2014.6 <- fitted(forecast(modArima,nrow(test.y),xreg=sig.x.test))[76:81]
data.2014 <- market_share_all[76:81,]
```

```{r}
(RMSE.ARIMA.with.Seasonality.6 <- sqrt(sum((data.2014 - fit.2014.6)^2)/6))
(MAE.ARIMA.with.Seasonality.6 <- sum(abs(data.2014 - fit.2014.6))/6)
(MAPE.ARIMA.with.Seasonality.6 <- 100*sum(abs((data.2014 - fit.2014.6)/fit.2014.6))/6)
```
Jane-June 2014:
```{r}
fit.2016.6 <- fitted(forecast(modArima,nrow(test.y),xreg=sig.x.test))[97:102]
data.2016 <- market_share_all[97:102,]
```

```{r}
(RMSE.ARIMA.with.Seasonality.6 <- sqrt(sum((data.2016 - fit.2016.6)^2)/6))
(MAE.ARIMA.with.Seasonality.6 <- sum(abs(data.2016 - fit.2016.6))/6)
(MAPE.ARIMA.with.Seasonality.6 <- 100*sum(abs((data.2016 - fit.2016.6)/fit.2016.6))/6)
```
May-Oct 2016:
```{r}
fit.2016.6 <- fitted(forecast(modArima,nrow(test.y),xreg=sig.x.test))[101:106]
data.2016 <- market_share_all[101:106,]
```

```{r}
(RMSE.ARIMA.with.Seasonality.6 <- sqrt(sum((data.2016 - fit.2016.6)^2)/6))
(MAE.ARIMA.with.Seasonality.6 <- sum(abs(data.2016 - fit.2016.6))/6)
(MAPE.ARIMA.with.Seasonality.6 <- 100*sum(abs((data.2016 - fit.2016.6)/fit.2016.6))/6)
```

```{r}
training.x[101:106,]
```



Fitting ARIMA without seasonality
Deseasonaling the dataset and plotting：
```{r}
decompose_data = decompose(share.train, "additive")
adjust_market_share = share.train - decompose_data$seasonal
plot(adjust_market_share)
```
Fit ARIMA to deseasoned data:
```{r}
modArima_desea <- auto.arima(adjust_market_share, xreg=sig.x.train)
modArima_desea
```
```{r}
Forecasted_values_des<-forecast(modArima_desea,nrow(test.y),xreg=sig.x.test)
Final_forecasted_values_des<-Forecasted_values_des$mean
(Final_forecasted_values_des)

(RMSE.ARIMA.without.Seasonality.6 <- sqrt(sum((test.y - Final_forecasted_values_des)^2)/6))
(MAE.ARIMA.without.Seasonality.6 <- sum(abs(test.y - Final_forecasted_values_des))/6)
(MAPE.ARIMA.without.Seasonality.6 <- 100*sum(abs((test.y - Final_forecasted_values_des)/Final_forecasted_values_des))/6)
```
Plotting forecast values after seasonality removal:
```{r}
forecast(modArima_desea,nrow(test.y),xreg=sig.x.test) -> fc1
ARIMA.without.Seasonality.plot.6 <- autoplot(market_share_all, series="Data") + 
autolayer(fc1, series="Forecast") + 
autolayer(fitted(fc1), series="Fitted")

ARIMA.without.Seasonality.plot.6 <- ARIMA.without.Seasonality.plot.6 + labs(title = "ARIMA.without.Seasonality.6") + labs(caption = "RMSE = 0.03067566, MAE = 0.02969591, MAPE = 13.53839")
ARIMA.without.Seasonality.plot.6
```
Plot only the foretasted period:
```{r}
forecast(modArima_desea,nrow(test.y),xreg=sig.x.test)$mean -> fc
ARIMA.Corr.without.Seasonality.forecast.6 <- autoplot(share.test, series="Data") + autolayer(fc, series="Forecast")

ARIMA.Corr.without.Seasonality.forecast.6 <- ARIMA.Corr.without.Seasonality.forecast.6 + labs(title = "ARIMA.Corr.without.Seasonality.forecast.6") + labs(caption = "RMSE = 0.03067566, MAE = 0.02969591, MAPE = 13.53839")
ARIMA.Corr.without.Seasonality.forecast.6
```
Use glm:
```{r}
glm.train <- glm(adjust_market_share~., data=as.data.frame(sig.x.train))
summary(glm.train)
(glm.prediction <- predict(glm.train, as.data.frame(sig.x.test)))

(RMSE.GLM.6 <- sqrt(sum((test.y - glm.prediction)^2)/6))
(MAE.GLM.6 <- sum(abs(test.y - glm.prediction))/6)
(MAPE.GLM.6 <- 100*sum(abs((test.y - glm.prediction)/glm.prediction))/6)
```

Plotting forecast values:
```{r}
fc1 <- ts(glm.prediction, start=c(2019.6),frequency=12)
fitted <- ts(glm.train$fitted.values, start=c(2008.1),frequency=12)
GLM.plot.6 <- autoplot(market_share_all, series="Data") + 
autolayer(fc1, series="Forecast") + 
autolayer(fitted, series="Fitted")

GLM.plot.6 <- GLM.plot.6 + labs(title = "GLM.6") + labs(caption = "RMSE =  0.05176947, MAE = 0.05613335, MAPE = 24.50887")
GLM.plot.6
```

Plot only the foretasted period:
```{r}
fc1 <- ts(glm.prediction, start=c(2019.6),frequency=12)
glm.forecast.6 <- autoplot(share.test, series="Data") + autolayer(fc1, series="Forecast")

glm.forecast.6 <- glm.forecast.6 + labs(title = "glm.forecast.6")+ labs(caption = "RMSE =  0.05176947, MAE = 0.05613335, MAPE = 24.50887")
glm.forecast.6
```