---
title: "ARIMA Baseline model"
author: "Kechen Zhao 957398"
date: "17/10/2021"
output: word_document
---

ARIMA 12months prediction modelling
```{r}
library(forecast)
library(lubridate)
library(tseries)
library(fpp)
library(ggplot2)
library(caret)
library(patchwork)
# import data
data <- read.csv("/Users/abigail/Dropbox/Data\ Science\ Project/Combined\ Datasets\ For\ AUS\ and\ US/08-19_combined.csv")
```
Construct training and test set:
```{r}
rownames(data) <- data$month
data <- subset(data, select = -month)

x <- subset(data, select = -import_ratio)
y <- subset(data, select = import_ratio)

market_share_all <- ts(y, start=c(2008.1),frequency=12)
```
```{r}
training.x <- x[1:132,]
training.y <- y[1:132,]
share.train <- ts(training.y, start=c(2008.1),frequency=12)
test.x <- x[133:144,]
test.y <- y[133:144,]
share.test <- ts(test.y, start=c(2019.1),frequency=12)
```
```{r}
Seasonaling <- plot(share.train)+title("Market Share with Seasonality")
```

```{r}
modArima <- auto.arima(share.train)
summary(modArima)
```

```{r}
Forecasted_values<-forecast(modArima, 12)
Final_forecasted_values<-Forecasted_values$mean
Final_forecasted_values
```
Calculating the MSE in the test dataset:
```{r}
test.y
```

```{r}
(RMSE.ARIMA.with.Seasonality.12 <- sqrt(sum((test.y - Final_forecasted_values)^2)/12))
(MAE.ARIMA.with.Seasonality.12 <- sum(abs(test.y - Final_forecasted_values))/12)
(MAPE.ARIMA.with.Seasonality.12 <- 100*sum(abs((test.y - Final_forecasted_values)/Final_forecasted_values))/12)
```
Plotting forecast values:
```{r}
forecast(modArima, 12) -> fc
ARIMA.with.Seasonality.plot.12 <- autoplot(market_share_all, series="Data") + 
autolayer(fc, series="Forecast") + 
autolayer(fitted(fc), series="Fitted")

ARIMA.with.Seasonality.plot.12 <- ARIMA.with.Seasonality.plot.12 + labs(title = "ARIMA.with.Seasonality.12.Baseline") + labs(caption = "RMSE = 0.02462555, MAE = 0.01989308, MAPE = 9.346342")
ARIMA.with.Seasonality.plot.12
```
Plot only the foretasted period:
```{r}
forecast(modArima,12)$mean -> fc
ARIMA.with.Seasonality.forecast.12 <- autoplot(share.test, series="Data") + autolayer(fc, series="Forecast")

ARIMA.with.Seasonality.forecast.12 <- ARIMA.with.Seasonality.forecast.12 + labs(title = "ARIMA.with.Seasonality.forecast.12.Baseline") + labs(caption = "RMSE = 0.02462555, MAE = 0.01989308, MAPE = 9.346342")
ARIMA.with.Seasonality.forecast.12
```
```{r}
ARIMA.with.Seasonality.plot.12 + ARIMA.with.Seasonality.forecast.12
```

Fitting ARIMA without seasonality
Deseasonaling the dataset and plotting：
```{r}
decompose_data = decompose(share.train, "additive")
adjust_market_share = share.train - decompose_data$seasonal
Deseasonaling <- plot(adjust_market_share) + title("Market Share without Seasonality")
```
```{r}
par(mfrow=c(1,2))
plot(share.train)+title("Market Share with Seasonality")
plot(adjust_market_share) + title("Market Share without Seasonality")
```

Fit ARIMA to deseasoned data:
```{r}
modArima_desea <- auto.arima(adjust_market_share)
modArima_desea
```

```{r}
Forecasted_values_des<-forecast(modArima_desea,12)
Final_forecasted_values_des<-Forecasted_values_des$mean
(Final_forecasted_values_des)

(RMSE.ARIMA.without.Seasonality.12 <- sqrt(sum((test.y - Final_forecasted_values_des)^2)/12))
(MAE.ARIMA.without.Seasonality.12 <- sum(abs(test.y - Final_forecasted_values_des))/12)
(MAPE.ARIMA.without.Seasonality.12 <- 100*sum(abs((test.y - Final_forecasted_values_des)/Final_forecasted_values_des))/12)
```

Plotting forecast values after seasonality removal:
```{r}
forecast(modArima_desea,12) -> fc1
ARIMA.without.Seasonality.plot.12 <- autoplot(market_share_all, series="Data") + 
autolayer(fc1, series="Forecast") + 
autolayer(fitted(fc1), series="Fitted")

ARIMA.without.Seasonality.plot.12 <- ARIMA.without.Seasonality.plot.12 + labs(title = "ARIMA.without.Seasonality.12.Baseline") + labs(caption = "RMSE = 0.02745205, MAE = 0.01913106, MAPE = 8.613308")
ARIMA.without.Seasonality.plot.12
```

Plot only the foretasted period:
```{r}
forecast(modArima_desea,12)$mean -> fc
ARIMA.without.Seasonality.forecast.12 <- autoplot(share.test, series="Data") + autolayer(fc, series="Forecast")

ARIMA.without.Seasonality.forecast.12 <- ARIMA.without.Seasonality.forecast.12 + labs(title = "ARIMA.without.Seasonality.forecast.12.Baseline") + labs(caption = "RMSE = 0.02745205, MAE = 0.01913106, MAPE = 8.613308")
ARIMA.without.Seasonality.forecast.12
```
```{r}
ARIMA.without.Seasonality.plot.12 + ARIMA.without.Seasonality.forecast.12
```




ARIMA 6months prediction modelling
```{r}
data <- read.csv("/Users/abigail/Dropbox/Data\ Science\ Project/Combined\ Datasets\ For\ AUS\ and\ US/08-19_combined.csv")
```
Construct training and test set:
```{r}
rownames(data) <- data$month
data <- subset(data, select = -month)

x <- subset(data, select = -import_ratio)
y <- subset(data, select = import_ratio)

market_share_all <- ts(y, start=c(2008.1),frequency=12)
```
```{r}
training.x <- x[1:138,]
training.y <- y[1:138,]
share.train <- ts(training.y, start=c(2008.1),frequency=12)
test.x <- x[139:144,]
test.y <- y[139:144,]
share.test <- ts(test.y, start=c(2019.6),frequency=12)
```
```{r}
modArima <- auto.arima(share.train)
summary(modArima)
```
```{r}
Forecasted_values<-forecast(modArima,6)
Final_forecasted_values<-Forecasted_values$mean
Final_forecasted_values
```
Calculating the MSE in the test dataset:
```{r}
(RMSE.ARIMA.with.Seasonality.6 <- sqrt(sum((test.y - Final_forecasted_values)^2)/6))
(MAE.ARIMA.with.Seasonality.6 <- sum(abs(test.y - Final_forecasted_values))/6)
(MAPE.ARIMA.with.Seasonality.6 <- 100*sum(abs((test.y - Final_forecasted_values)/Final_forecasted_values))/6)
```
Plotting forecast values:
```{r}
forecast(modArima,6) -> fc
ARIMA.with.Seasonality.plot.6 <- autoplot(market_share_all, series="Data") + 
autolayer(fc, series="Forecast") + 
autolayer(fitted(fc), series="Fitted")

ARIMA.with.Seasonality.plot.6 <- ARIMA.with.Seasonality.plot.6 + labs(title = "ARIMA.with.Seasonality.6.Baseline") + labs(caption = "RMSE = 0.02037986, MAE = 0.01646762, MAPE = 7.04457")
ARIMA.with.Seasonality.plot.6
```
Plot only the foretasted period:
```{r}
forecast(modArima,6)$mean -> fc
ARIMA.with.Seasonality.forecast.6 <- autoplot(share.test, series="Data") + autolayer(fc, series="Forecast")

ARIMA.with.Seasonality.forecast.6 <- ARIMA.with.Seasonality.forecast.6 + labs(title = "ARIMA.with.Seasonality.forecast.6.Baseline") + labs(caption = "RMSE = 0.02037986, MAE = 0.01646762, MAPE = 7.04457")
ARIMA.with.Seasonality.forecast.6
```
```{r}
ARIMA.with.Seasonality.plot.6 + ARIMA.with.Seasonality.forecast.6
```
Fitting ARIMA without seasonality
Deseasonaling the dataset and plotting：
```{r}
decompose_data = decompose(share.train, "additive")
adjust_market_share = share.train - decompose_data$seasonal
plot(adjust_market_share)
```
Fit ARIMA to deseasoned data:
```{r}
modArima_desea <- auto.arima(adjust_market_share)
modArima_desea
```
```{r}
Forecasted_values_des<-forecast(modArima_desea,6)
Final_forecasted_values_des<-Forecasted_values_des$mean
(Final_forecasted_values_des)

(RMSE.ARIMA.without.Seasonality.6 <- sqrt(sum((test.y - Final_forecasted_values_des)^2)/6))
(MAE.ARIMA.without.Seasonality.6 <- sum(abs(test.y - Final_forecasted_values_des))/6)
(MAPE.ARIMA.without.Seasonality.6 <- 100*sum(abs((test.y - Final_forecasted_values_des)/Final_forecasted_values_des))/6)
```
Plotting forecast values after seasonality removal:
```{r}
forecast(modArima_desea,5) -> fc1
ARIMA.without.Seasonality.plot.6 <- autoplot(market_share_all, series="Data") + 
autolayer(fc1, series="Forecast") + 
autolayer(fitted(fc1), series="Fitted")

ARIMA.without.Seasonality.plot.6 <- ARIMA.without.Seasonality.plot.6 + labs(title = "ARIMA.without.Seasonality.6.Baseline") + labs(caption = "RMSE = 0.01558521, MAE = 0.01393499, MAPE = 5.965604")
ARIMA.without.Seasonality.plot.6
```
Plot only the foretasted period:
```{r}
forecast(modArima_desea,6)$mean -> fc
ARIMA.without.Seasonality.forecast.6 <- autoplot(share.test, series="Data") + autolayer(fc, series="Forecast")

ARIMA.without.Seasonality.forecast.6 <- ARIMA.without.Seasonality.forecast.6 + labs(title = "ARIMA.without.Seasonality.forecast.6.Baseline") + labs(caption = "RMSE = 0.01558521, MAE = 0.01393499, MAPE = 5.965604")
ARIMA.without.Seasonality.forecast.6
```
```{r}
ARIMA.without.Seasonality.plot.6 + ARIMA.without.Seasonality.forecast.6
```



















